The importance of reproducibility in COVID-19 research: the case of population density and the spread of the pandemic

### 2021-05-19 R0-Density-Reanalysis/R0-Density-Reanalysis.Rmd
Added reference to butress the point of risk perception

A plausible behavioral adaptiation during a pandemic, especially one broadcast as widely and intensely as COVID-19, is risk compensation. Risk compensation is a process whereby people adjust their behavior in response to their _perception_ of risk [@Noland1995perceived; @Richens2000condoms; @Phillips2011risk]. For example, it is possible that people who listened to the message of leaders saying that they were safe because of low density may not have taken adequate precautions. People in dense places who could more directly observe the impact of the pandemic may have become overly cautious. Both Paez et al. [-@Paez2020spatio] and Hamidi et al. [-@Hamidi2020density] posit this mechanism (i.e., greater compliance with social distancing in denser regions) to explain the results of their analyses. The evidence available does indeed show that there were important changes in behavior during the pandemic, at least with respect to mobility [@Jamal2020Changes; @Harris2021Changes; @Molloy2020Tracing]; furthermore, shelter in place orders may have had greater buy-in from the public in higher density regions [@Feyman2020effectiveness], and the behavior may have persisted beyond the duration of official social-distancing policies [@Praharaj2020Using]. In addition, there is evidence that changes in mobility correlated with the trajectory of the pandemic [@Paez2020using; @Noland2021mobility]. Given the potential for behavioral adaptation, the question of density becomes more nuanced: it is not just a matter of proximity, but also of human behavior, which is better studied using population-level data and models.

>>>>

A plausible behavioral adaptation during a pandemic, especially one broadcast as widely and intensely as COVID-19, is risk compensation. Risk compensation is a process whereby people adjust their behavior in response to their _perception_ of risk [@Noland1995perceived; @Richens2000condoms; @Phillips2011risk]. In the case of COVID-19, Chauhan et al. [@Chauhan2021covid] have found that perception of risks in the US varies between rural, suburban, and urban residents, with rural residents in general displaying less concerns about the virus. It is possible that people who listened to the message of leaders saying that they were safe because of low density may not have taken adequate precautions against the virus. People in dense places who could more directly observe the impact of the pandemic may have become overly cautious. Both Paez et al. [-@Paez2020spatio] and Hamidi et al. [-@Hamidi2020density] posit this mechanism (i.e., greater compliance with social distancing in denser regions) to explain the results of their analyses. The evidence available does indeed show that there were important changes in behavior during the pandemic, at least with respect to mobility [@Jamal2020Changes; @Harris2021Changes; @Molloy2020Tracing]; furthermore, shelter in place orders may have had greater buy-in from the public in higher density regions [@Feyman2020effectiveness], and the behavior may have persisted beyond the duration of official social-distancing policies [@Praharaj2020Using]. In addition, there is evidence that changes in mobility correlated with the trajectory of the pandemic [@Paez2020using; @Noland2021mobility]. Given the potential for behavioral adaptation, the question of density becomes more nuanced: it is not just a matter of proximity, but also of human behavior, which is better studied using population-level data and models.
